---
title: "Unsupervised_learning_2"
author: "Joan M. Valls Cuevas"
date: "5/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Unsupervised learning 2

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
```
Q3
```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
```
```{r}

length(grep("_mean", colnames(wisc.df)))


```


```{r}
row.names(wisc.data) <- wisc.df$id
```

```{r}
diagnosis <- wisc.df$diagnosis

```
Check the mean of the columns in order to see if scaling is appropriate
```{r}
round(colMeans(wisc.data), 1)
#apply(wisc.data, 1 for row 2 for col, SD or mean etc)
```



Create a PCA plot and scaling as appropriate
```{r}
wisc.pr <- prcomp(wisc.data, scale. = TRUE)
```

Summary of the PCA data
```{r}
summary(wisc.pr)
```

Creating a biplot of data
```{r}
biplot(wisc.pr)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = diagnosis)
```


```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3], col = diagnosis)
```

```{r}
pr.var <- wisc.pr$sdev^2

pve <- pr.var/sum(pr.var)

plot(pve, type = "o", xlab = "Principal Component", ylab = "Proportion of Variance explained")
```

```{r}
barplot(pve*100, xlab = "Percent of Variance explained", names.arg = paste0("PC", 1:30), cex.names = 0.5, axis.lty = 1)

sum(pr.var)
```



```{r}
data.scaled <- scale(wisc.data)

data.dist <- dist(data.scaled, method = "euclidean")

wisc.hclust <- hclust(data.dist)

plot(wisc.hclust)


```

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
```



```{r}
table(wisc.hclust.clusters, diagnosis)
```


##Section 5. combining methods (PCA + clustering)

I am going to start with principal components that capture 90% or more of the origical dataset (PC1 to PC7)
```{r}

wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), method = "ward.D2")
```
Plot the new hierarchical cluster
```{r}
plot(wisc.pr.hclust)

grps <- cutree(wisc.pr.hclust, k = 2)

table(grps)

table(grps, diagnosis)
plot(wisc.pr$x[,1:2], col = grps )

```

##Section 7: Predicting using new data

```{r}

url <- "https://tinyurl.com/new-samples-CSV"

new <- read.csv(url)

npc <- predict(wisc.pr, newdata=new)
npc

plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")

```


```{r}
sessionInfo()
```

