---
title: "Class 14: Genome informatics part 2"
author: "Joan M. Valls Cuevas"
date: "5/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## section 1

```{r}
metadata <- read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)


```

Q1. How many genes are we dealing with?
```{r}
nrow(counts)
```

Q2.
```{r}


all(colnames(counts)[-1] == metadata$id)

```


Let's access just the controls conts data so we can calculate an averae counts value for each gene in these control samples
```{r}
control <- metadata[metadata[,"dex"]=="control",]
control
#now we want to get the cols in counts with these 
control.mean <- rowSums( counts[ ,control$id] )/length(control$id) 
names(control.mean) <- counts$ensgene
head(control.mean)
```



```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.mean <- rowSums( counts[ ,treated$id] )/ length(treated$id)
names(treated.mean) <- counts$ensgene

mean.counts <- data.frame(control.mean,treated.mean)



```


```{r}
plot(mean.counts)
```



```{r}
plot(mean.counts, log="xy")
```

```{r}
mean.counts$log2fc <- log2((mean.counts[,"treated.mean"])/mean.counts[,"control.mean"])
head(mean.counts)

```


```{r}
#select the genes that are zero
to.rm <- unique(which(mean.counts[,1:2] == 0, arr.ind = TRUE )[,1])
#remove these rows as we can't really say much about these
mycounts <- mean.counts[-to.rm]

```

```{r}

pos <- which(mean.counts[,3] >=2, arr.ind = TRUE)

pos_count <- mean.counts[pos,]

pos_count <- pos_count[!is.infinite(rowSums(pos_count)),]
nrow(pos_count)
```



```{r}
neg <- which(mean.counts[,3] <=-2, arr.ind = TRUE)

neg_count <- mean.counts[neg,]
neg_count <- neg_count[!is.infinite(rowSums(neg_count)),]
nrow(neg_count)
```

## 4. Adding annotation data

```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```




using the **merge()** fucntion to add another data to our 'mycounts' results
```{r}
mycounts.anno <- merge(x= mycounts, y = anno, by.x = 0, by.y = "ensgene")
```


## We can use bioconductor to do this as well... 

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```



```{r}
columns(org.Hs.eg.db)
```




```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts), # Our genenames
                     keytype="ENSEMBL",        # The format of our genenames
                     column="SYMBOL",          # The new format we want to add
                     multiVals="first")
```





```{r}
mycounts$entrez <- mapIds(org.Hs.eg.db,
                          keys = row.names(mycounts),
                          keytype ="ENSEMBL",
                          column = "ENTREZID",
                          multiVals = "first")


mycounts$uniprot <- mapIds(org.Hs.eg.db,
                          keys = row.names(mycounts),
                          keytype ="ENSEMBL",
                          column = "UNIPROT",
                          multiVals = "first")
```


```{r}
citation("DESeq2")
```


Building our deseq dataset object
```{r}
library(DESeq2)
dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = metadata,
                              design = ~dex,
                              tidy = TRUE)
```

```{r}
dds <- DESeq(dds)
```


Get our results
```{r}
res <- results(dds)
res
```


```{r}
summary(res)
```

```{r}
res05 <- results(dds, alpha= 0.05)
summary(res05)
```

```{r}
#coloring and building my graph
mycols <- rep("gray", nrow(res))
mycols[res$padj < 0.05] <- "red"
mycols[abs(res$log2FoldChange) > 2 & res$padj < 0.05] <- "blue"
plot(res$log2FoldChange, -log(res$padj), col = mycols)
abline(h= -log(0.05), v = c(2,-2), lty = 2, col = "darkgray")
```

An alternative way to make volcano plots is with the Enhanced

But first lets put proper gene names on our results (i.e. gene 'symbols')

```{r}
x <- as.data.frame(res)
x$symbol <- mapIds(org.Hs.eg.db, 
                   keys=row.names(x),
                   keytype="ENSEMBL",
                   column="SYMBOL",
                   multiVals="first")
```

```{r}
library(ggplot2)
library(ggrepel)
library(EnhancedVolcano)
```







