---
title: "Class 7: R functions and packages"
author: "Joan M. Valls Cuevas"
date: "4/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## More on function writting

First we will revisit out function from the last lecture
```{r}
source("http://tinyurl.com/rescale-R")
```

Test the  **rescale()** function

```{r}
rescale(1:10)
```
Test **rescale()** using a string
```{r}
##rescale(c(1:10, "string"))
```
Now that we see an error was thrown to us, we will use a **warning()** or **stop()** function. use the rescale2() function which applies this
```{r}
##rescale2(c(1:10, "string"))

```

##Function practice

Write a function to identify NA elements in two vectors

Starting with a simple example to test body of function
```{r}
x <- c(1,2,NA,3,NA)
y <- c(NA, 3, NA, 3, 4)
is.na(x)
is.na(x)
```

Test the logical AND operator to see if it detects both NA values

```{r}
is.na(x) & is.na(y)
```

Find number of TRUE statements that are both TRUE
```{r}
sum(is.na(x) & is.na(y))
```

Using code snippet to make function
```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
} 
```
apply function that is made
```{r}
both_na(x,y)
```
second test case

```{r}
both_na(c(NA,NA,NA), c(1,NA,NA,NA))
```
Need to check if length is the same of both vector.

```{r}
x <- c(NA,NA,NA)
y <- c(1,NA,NA,NA)
length(x) != length(y)

```
Try the both_na3() function with extra features

```{r}
both_na3 <- function(x, y) {

if(length(x) != length(y)) {
stop("Input x and y should be vectors of the same length")

}
  na.in.both <- ( is.na(x) & is.na(y) )
  na.number  <- sum(na.in.both)
  na.which   <- which(na.in.both)
  message("Found ", na.number, " NA's at position(s):",
          paste(na.which, collapse=", ") )
  return( list(number=na.number, which=na.which) )
}
```

#Write function to grade homework
start with a simple example
```{r}
#student 1
stu1 <- c(100, 100, 100, 100, 100, NA ,100, 90)
#student 2
stu2 <- c(100,NA,90,90,90,90,87,80)

min(stu1)
which.min(stu1)
NA_val <- NA
stu1[which.min(stu1)] <- NA_val
stu1


```

```{r}
grade <- function(x) {

(sum(x, na.rm = TRUE) - min(x ,  na.rm = TRUE)) / (length(x) - 1)


}
```

```{r}
grade(stu1)
```

```{r}
url <- "https://tinyurl.com/gradeinput"

students <- read.csv(url, row.names = 1)

ans <- apply(students, 1, grade)
```


```{r}
sort(ans, decreasing = TRUE)
```

##one last function example

Find the intersection of two vectors

```{r}
input <-  source("http://tinyurl.com/rescale-R")

```
```{r}
x <- df1$IDs
y <- df2$IDs

intersect(x,y)
```

```{r}
x
y

x %in% y
x[x %in% y]

```






```{r}
gene_intersect <- function(x, y) {
  cbind(x[x %in% y],
        y[y %in% x])
  
}

```

```{r}
merge(df1, df2, by="IDs")
```










