---
title: "Class 12: structural informatics (prt2)"
author: "Joan M. Valls Cuevas"
date: "5/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library("bio3d")

file.name <- get.pdb("1hsg")

hiv <- read.pdb(file.name)
hiv
```

```{r}
#
prot <- atom.select(hiv, "protein", value = TRUE)
prot
lig <- atom.select(hiv, "ligand", value = TRUE)
lig


```


```{r}
write.pdb(prot, file = "1hsg_protein.pdb")
write.pdb(lig, file = "1hsg_ligand.pdb")
```


##Add hydrogens and charges in ADT

We opened our protein only PDB file in AutoDocTools and added hydrogens and atom-types needed for docking calculations.

#Run docking 

We will use the AutoDoc VIna here at the UNIX command line

The command I used was:
~/Downloads/autodock_vina_1_1_2_mac/bin/vina --config config.txt --log log.txt



## Process results back in R

```{r}
res <- read.pdb("all.pdbqt", multi = TRUE)
res
write.pdb(res, file = "results.pdb")
```



Compare out docking poses to the MERK drug 
```{r}

ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)

  
```

Scenario two
NMA in bio3d

```{r}
pdb <- read.pdb("1hel")
modes <- nma(pdb)
m7 <- mktrj(modes, mode=7, file="nma_7.pdb")
#Shows where molecule is most likely to be flexible
plot(modes, sse=pdb)
```


















